---
title: "CompBio_LectureNotes_03212017"
author: "Emily Mikucki"
date: "3/21/2017"
output: html_document
---

###**Guest Lecture: Brandon Ogbunu**   
* Darwin's Dangerous Idea --> Dennett's Calculator (Wrongulator); doesn't mask ignaorance, is a means to unmasking ignorance   
* Modelling = understanding the pieces that will help you understand the larger picture 

####**Mathematical model**   
* focuses on specific quantitative features of hte scenario, ignores others   
* mathematical description of a scenario or situation from the real-world   
* simplification, abstraction
* involves hypotheses that can be tested against real data and refined if desired 
* improved understanding of real-world scenario 

####**Why study epidemic models?**   
* supplement statistical extrapolations   
* learn more about the qualitative dynamics of a disease   
* test hypotheses about (for example) prevention strategies, disease transmission, significant charactersitics, etc.   

####**Different kinds of models:**  
* deterministic   
* discrete-time   
* stochastic   
* individual-based model (negative: if population is large you get "average" anyways...)   
* network models   

####**SIR epidemic model**    
* consider a disease spread by contact with infected individuals   
* define rate of change (susceptible, infected, recovered...)   

####**Ordinary differential equations**   
* pedator prey models   

####**Basic differential equation model in R**   
* Use EpiModel   

####Basic SI Model: Starting with a basic Susceptible-Infected (SI) disease model in which there is random mixing within the population
```{r}
#install.packages("EpiModel", dependencies = TRUE)
library(EpiModel)
param <- param.dcm(inf.prob = 0.2, act.rate = 0.25)
init <- init.dcm(s.num = 500, i.num = 1) #sets the transmission probability per act
control <- control.dcm(type = "SI", nsteps = 500) #collects other structural model controls like the model type and number of time steps for the simulation

mod <- dcm(param, init, control)
mod
plot(mod)
summary(mod, at = 150) #Determines num susceptible and num infected @ time 150 
```

####SIR Model with Demography: In a Susceptible-Infectious-Recovered (SIR) model, infected individuals transition from disease into a life-long recovered state in which they are never again susceptible

```{r}
#parameters for the recovery rate, birth rate, and state-specific death rates are entered
param <- param.dcm(inf.prob = 0.2, act.rate = 1, rec.rate = 1/20,
                   b.rate = 1/95, ds.rate = 1/100, di.rate = 1/80, dr.rate = 1/100)
init <- init.dcm(s.num = 1000, i.num = 1, r.num = 0) #specify the number of initially recovered
control <- control.dcm(type = "SIR", nsteps = 500, dt = 0.5) #obtain model results in fractional time units
mod <- dcm(param, init, control)

par(mar = c(3.2, 3, 2, 1), mgp = c(2, 1, 0), mfrow = c(1, 2)) #used to change some default graphical options
plot(mod, popfrac = FALSE, alpha = 0.5,
     lwd = 4, main = "Compartment Sizes")
plot(mod, y = "si.flow", lwd = 4, col = "firebrick",
     main = "Disease Incidence", leg = "n") #This is really cool! Plot the prevalences for all compartments in the model, but in the right plot we override that using the y argument to specify that disease incidence

par(mfrow = c(1, 1))
comp_plot(mod, at = 50, digits = 1) #The plot shows the three state sizes and flows at t50t50.
```

####Change the Parameters of SIR Model with Demography : 

```{r}
param <- param.dcm(inf.prob = 0.5, act.rate = 1, rec.rate = 5/20,
                   b.rate = 5/95, ds.rate = 5/100, di.rate = 5/80, dr.rate = 5/100)
init <- init.dcm(s.num = 1000, i.num = 1, r.num = 0)
control <- control.dcm(type = "SIR", nsteps = 500, dt = 0.5)
 mod <- dcm(param, init, control)
mod
plot(mod)
summary(mod, at = 150)
par(mar = c(3.2, 3, 2, 1), mgp = c(2, 1, 0), mfrow = c(1, 2)) #used to change some default graphical options
plot(mod, popfrac = FALSE, alpha = 0.5,
     lwd = 4, main = "Compartment Sizes")
plot(mod, y = "si.flow", lwd = 4, col = "firebrick",
     main = "Disease Incidence", leg = "n")
```
