---
title: "First Markdown Trial"
author: "Emily Mikucki"
date: "January 24-26, 2017"
output:
  html_document: default
  pdf_document: default
---

This text has *italics*.  

This text is **bold**

# First Header

## Second Header

### etc. etc. 

these         words are widely        spaced

One line   
Next line   
Third line   

(add 3 spaces after to make lines)

- first item
- second item
    * indented item
    * another indent 
    
[linked text](website address)

footnote at the bottom of the page^[This will magically appear at the bottom with active links in both directions!]

> quoted text is formatted this way 

First Header  | Second Header
------------- | -------------
Content Cell  | Content Cell
Content Cell  | Content Cell

`plain text` shows this way

```
now     I     can space this!
More of my     *text here*!
inside this text none of the commands work
```

in line r code `r 3+ pi`

```{r}
# Annotation/comment
Time <- seq(1,10) # seq. makes integer sequence
print(Time)   # shows contents
Resp <- runif(10) #random uniform (runif), gives 10 random uniforms
print(Resp)
plot(x=Time, y=Resp, type="b") #b is both  point and line connecting 

```

### In line equation 
Put an equation $a = b + c$ in my sentence 

Set equation on its own line with this 

$$ a = b + c$$

###Subscripts in LaTex

$$H_0 = Z_{a + b}$$

In Markdown, we say H~0~

###Superscripts 

$$S = cA^z$$

###Combining elements

$$S=cA^z_1 = z_{2+ x_0}$$

###Special symbols
$$\alpha = \frac{\beta}{\gamma + \delta_x}$$

###Using the backslash 
$$\backslash \alpha \le b \backslash$$

###Summation sign

$$z = \sum_{i=1}^X{K}$$

###Remndering plaintext in LaTex

$$P(\mbox{Occurence Of Species A}) = Z$$


```{r}
plot(runif(20))
print(seq(1,5))
```

